---
title: "R Notebook"
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
---

Loading the Package:


```{r}
library(plyr)
library(dplyr)
library(ggplot2)


#Loading the Data
dataNW<-read.csv("Demand Forecasting dataset.csv")
head(dataNW)



```


Data Explorationg:

1.Checking all the Variables

```{r}
library(DataExplorer)
plot_str(dataNW)
```

2. Checking Missing Values

```{r}
plot_missing(dataNW)

```

```{r}
str(dataNW)

```




3. Changing Order date and Shipping Date into "Date" format




```{r}
library(lubridate)
dataNW$Order.Date<-as.Date(dataNW$Order.Date,'%m/%d/%Y')
dataNW$Ship.Date<-as.Date(dataNW$Ship.Date,'%m/%d/%Y')
```


4. Data Visualization to check the Highest Sales with respect to Segment

```{r}
segment_sale = aggregate(Sales ~ Segment, data = dataNW, sum)
segment_sale
```



```{r}
ggplot(segment_sale, aes(x = Segment, y = Sales)) + geom_bar(stat =
"identity", fill = 'khaki', colour = 'black') + ggtitle("Sales by Segment") + labs(y = "Segment", x =
"Sales")
```


We can that Sales in Consumer Sector is Highest








Time  Series Forecasting 

1. Segment- Counsumer:


Extracting the Order Date and Sales filtered with Segment- Consumer.

```{r}

library(dplyr)
Consumer<-dataNW %>% 
  select(Order.Date, Sales) %>%
  filter(dataNW$Segment == "Consumer")

```


Displaying the data order by Order_Date:

```{r}
counsumer<-Consumer[order(as.Date(Consumer$Order.Date, "%m/%d/%Y"), decreasing = FALSE),] 
tail(counsumer)
```


Adding Year and Month

```{r}
counsumer$year <- format(as.Date(counsumer$Order.Date), "%Y")
counsumer$month <- format(as.Date(counsumer$Order.Date), "%m")

head(counsumer)
```



Using SQL Package to select year and month and the total sales order by year and month

```{r}
library(sqldf)
counsumer<-sqldf("select year, month, sum(Sales) as Month_Sales from counsumer group by year, month order by year, month")
tail(counsumer)

```
Analysis- Total sum of Sales in December 2017 is $50232.46


Changing the data into time series-

```{r}
timeSeriesCounsumer<-ts(counsumer$Month_Sales, frequency=12,start=c(2014,1),end=c(2017,12))
plot(timeSeriesCounsumer)
head(timeSeriesCounsumer)
```

There is Seasonality on the time series

```{r}
decompose_consumer<-decompose(timeSeriesCounsumer)
plot(decompose_consumer)
```


TIME SERIES FORCASTING

Simple Smoothing

For next 2 years:

```{r}
library(forecast)
simple_smoothing1 <- forecast(timeSeriesCounsumer)
summary(simple_smoothing1)

```

December 2017 Total Sale in December is 50232.46

After 1 year, forecasted sale would be 44695.30



50232.46-44695.30/50232.46= 11% 




```{r}
accuracy(simple_smoothing1)
```

```{r}
plot(simple_smoothing1)
```


Plot Between Actual and Residual:

```{r}
library(ggplot2)
qplot(y = simple_smoothing1$residuals, x = timeSeriesCounsumer,
      ylab = "Residuals", xlab = "Actual values", 
      main = " Residuals vs. Actual plot") +
  stat_smooth(method = "loess", span = 0.1, colour = I("blue"), se = FALSE)
```


Since the sales of Cousmer might fall down 1 year later- Lets check the forecast of Cosunmer in the three 3 Regions :

FOR SOUTH

```{r}
library(dplyr)
Consumer_South<-dataNW %>% 
  select(Order.Date, Sales) %>%
  filter(dataNW$Segment == "Consumer",dataNW$Region=="South")
```


```{r}
tail(Consumer_South)
```

```{r}
Consumer_South<-Consumer_South[order(as.Date(Consumer_South$Order.Date, "%m/%d/%Y"), decreasing = FALSE),] 
tail(Consumer_South)
```

```{r}
Consumer_South$year <- format(as.Date(Consumer_South$Order.Date), "%Y")
Consumer_South$month <- format(as.Date(Consumer_South$Order.Date), "%m")

head(Consumer_South)
```



```{r}
library(sqldf)
Consumer_South<-sqldf("select year, month, sum(Sales) as Month_Sales from Consumer_South group by year, month order by year, month")
tail(Consumer_South)

```

Sum of Total Sales for Counsumer Segment in South in December 2017 is $11159.3222





```{r}
timeSeriesCounsumer_South<-ts(Consumer_South$Month_Sales, frequency=12,start=c(2014,1),end=c(2017,12))
plot(timeSeriesCounsumer_South)
```

```{r}
decompose_consumer_south<-decompose(timeSeriesCounsumer_South)
plot(decompose_consumer_south)
```


TIME SERIES FORCASTING

Simple Smoothing

For next 2 years:

```{r}
library(forecast)
simple_smoothing1_south <- forecast(timeSeriesCounsumer_South)
summary(simple_smoothing1_south)

```

After 1 year forecast of total sales would be $6525.516




2. For Central

```{r}
library(dplyr)
Consumer_Central<-dataNW %>% 
  select(Order.Date, Sales) %>%
  filter(dataNW$Segment == "Consumer",dataNW$Region=="Central")
```


```{r}
tail(Consumer_Central)
```

```{r}
Consumer_Central<-Consumer_Central[order(as.Date(Consumer_Central$Order.Date, "%m/%d/%Y"), decreasing = FALSE),] 
tail(Consumer_Central)
```

```{r}
Consumer_Central$year <- format(as.Date(Consumer_Central$Order.Date), "%Y")
Consumer_Central$month <- format(as.Date(Consumer_Central$Order.Date), "%m")

head(Consumer_Central)
```



```{r}
library(sqldf)
Consumer_Central<-sqldf("select year, month, sum(Sales) as Month_Sales from Consumer_Central group by year, month order by year, month")
tail(Consumer_Central)

```
Total Sum of Sales in December 2017 is $13793.727 for Counsmer Segment in Central region




```{r}
timeSeriesCounsumer_central<-ts(Consumer_Central$Month_Sales, frequency=12,start=c(2014,1),end=c(2017,12))
plot(timeSeriesCounsumer_central)
```

```{r}
decompose_consumer_central<-decompose(timeSeriesCounsumer_central)
plot(decompose_consumer_central)
```
There is no trend but there is sesonality.



TIME SERIES FORCASTING

Simple Smoothing

For next 2 years:

```{r}
library(forecast)
simple_smoothing1_central <- forecast(timeSeriesCounsumer_central)
summary(simple_smoothing1_central)

```
Forecast of Total sales after 1 year would be $9472.7526


3. FOR WEST


```{r}
library(dplyr)
Consumer_West<-dataNW %>% 
  select(Order.Date, Sales) %>%
  filter(dataNW$Segment == "Consumer",dataNW$Region=="West")
```


```{r}
tail(Consumer_West)
```

```{r}
Consumer_West<-Consumer_West[order(as.Date(Consumer_West$Order.Date, "%m/%d/%Y"), decreasing = FALSE),] 
tail(Consumer_West)
```

```{r}
Consumer_West$year <- format(as.Date(Consumer_West$Order.Date), "%Y")
Consumer_West$month <- format(as.Date(Consumer_West$Order.Date), "%m")

head(Consumer_West)
```



```{r}
library(sqldf)
Consumer_West<-sqldf("select year, month, sum(Sales) as Month_Sales from Consumer_West group by year, month order by year, month")
tail(Consumer_West)

```
Total Sales in December 2017 is $12589.989



```{r}
timeSeriesCounsumer_west<-ts(Consumer_West$Month_Sales, frequency=12,start=c(2014,1),end=c(2017,12))
plot(timeSeriesCounsumer_west)
```

```{r}
decompose_consumer_west<-decompose(timeSeriesCounsumer_west)
plot(decompose_consumer_west)
```


TIME SERIES FORCASTING

Simple Smoothing

For next 2 years:

```{r}
library(forecast)
simple_smoothing1_west <- forecast(timeSeriesCounsumer_west)
summary(simple_smoothing1_west)

```

Forecast of Total Sales after 1 year would be $11603.561



4. FOR EAST

```{r}
library(dplyr)
Consumer_East<-dataNW %>% 
  select(Order.Date, Sales) %>%
  filter(dataNW$Segment == "Consumer",dataNW$Region=="East")
```


```{r}
tail(Consumer_East)
```

```{r}
Consumer_East<-Consumer_East[order(as.Date(Consumer_East$Order.Date, "%m/%d/%Y"), decreasing = FALSE),] 
tail(Consumer_East)
```

```{r}
Consumer_East$year <- format(as.Date(Consumer_East$Order.Date), "%Y")
Consumer_East$month <- format(as.Date(Consumer_East$Order.Date), "%m")

head(Consumer_East)
```



```{r}
library(sqldf)
Consumer_East<-sqldf("select year, month, sum(Sales) as Month_Sales from Consumer_East group by year, month order by year, month")
tail(Consumer_East)

```
Total Sales in Decemeber 2017 is $12689.418


```{r}
timeSeriesCounsumer_east<-ts(Consumer_East$Month_Sales, frequency=12,start=c(2014,1),end=c(2017,12))
plot(timeSeriesCounsumer_east)
```

```{r}
decompose_consumer_east<-decompose(timeSeriesCounsumer_east)
plot(decompose_consumer_east)
```


TIME SERIES FORCASTING

Simple Smoothing

For next 2 years:

```{r}
library(forecast)
simple_smoothing1_central <- forecast(timeSeriesCounsumer_east)
summary(simple_smoothing1_central)

```

Forecast of Total Sales after 1 year would be $13327.104

Sale in the Central Region is going high.







CORPORATE


```{r}

library(dplyr)
Corporate<-dataNW %>% 
  select(Order.Date, Sales) %>%
  filter(dataNW$Segment == "Corporate")

```


```{r}
Corporate<-Corporate[order(as.Date(Corporate$Order.Date, "%m/%d/%Y"), decreasing = FALSE),] 
tail(Corporate)
```


```{r}
Corporate$year <- format(as.Date(Corporate$Order.Date), "%Y")
Corporate$month <- format(as.Date(Corporate$Order.Date), "%m")

head(Corporate)
```

```{r}
library(sqldf)
Corporate<-sqldf("select year, month, sum(Sales) as Month_Sales from Corporate group by year, month order by year, month")
tail(Corporate)

```


```{r}
timeSeriesCorporate<-ts(Corporate$Month_Sales, frequency=12,start=c(2014,1),end=c(2017,12))
plot(timeSeriesCorporate)
head(timeSeriesCorporate)
```

```{r}
timeSeriesCorporate_decompose<-decompose(timeSeriesCorporate)
plot(timeSeriesCorporate_decompose)
```


```{r}
simple_smoothing2 <- forecast(timeSeriesCorporate)
summary(simple_smoothing2)

```


30433.043


20524.43


```{r}
accuracy(simple_smoothing2)
```

```{r}
plot(simple_smoothing2)
```

```{r}
library(ggplot2)
qplot(y = simple_smoothing2$residuals, x = timeSeriesCorporate,
      ylab = "Residuals", xlab = "Actual values", 
      main = " Residuals vs. Actual plot") +
  stat_smooth(method = "loess", span = 0.1, colour = I("blue"), se = FALSE)
```

3. Home Office




```{r}

library(dplyr)
Home_Office<-dataNW %>% 
  select(Order.Date, Sales) %>%
  filter(dataNW$Segment == "Home Office")

```


```{r}
Home_Office<-Home_Office[order(as.Date(Home_Office$Order.Date, "%m/%d/%Y"), decreasing = FALSE),] 
tail(Home_Office)
```


```{r}
Home_Office$year <- format(as.Date(Home_Office$Order.Date), "%Y")
Home_Office$month <- format(as.Date(Home_Office$Order.Date), "%m")

head(Home_Office)
```

```{r}
library(sqldf)
Home_Office<-sqldf("select year, month, sum(Sales) as Month_Sales from Home_Office group by year, month order by year, month")
tail(Home_Office)

```




```{r}
timeSeriesHome_Office<-ts(Home_Office$Month_Sales, frequency=12,start=c(2014,1),end=c(2017,12))
plot(timeSeriesHome_Office)
head(timeSeriesHome_Office)
```

```{r}
timeSeriesHome_Office_decompose<-decompose(timeSeriesHome_Office)
plot<-timeSeriesHome_Office_decompose

```

```{r}
plot(timeSeriesHome_Office_decompose)

```


```{r}
simple_smoothing3 <- forecast(timeSeriesHome_Office)
summary(simple_smoothing3)

```


```{r}
plot(simple_smoothing3)
```

```{r}
library(ggplot2)
qplot(y = simple_smoothing3$residuals, x = timeSeriesHome_Office,
      ylab = "Residuals", xlab = "Actual values", 
      main = " Residuals vs. Actual plot") +
  stat_smooth(method = "loess", span = 0.1, colour = I("blue"), se = FALSE)
```

